---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---
```{r pkg-knitr-opts, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, fig.retina=2, message=FALSE, warning=FALSE)
options(width=60)
```

[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/udpprobe.svg?branch=master)](https://travis-ci.org/hrbrmstr/udpprobe) 
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/hrbrmstr/udpprobe?branch=master&svg=true)](https://ci.appveyor.com/project/hrbrmstr/udpprobe) 
[![Coverage Status](https://codecov.io/gh/hrbrmstr/udpprobe/branch/master/graph/badge.svg)](https://codecov.io/gh/hrbrmstr/udpprobe)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/udpprobe)](https://cran.r-project.org/package=udpprobe)

# udpprobe

Send User Datagram Protocol ('UDP') Probes and Receive Responses

## Description

Unfortunately (but understandably) R has no build in connection  support for
the user datagram protocol ('UDP'). We aim to somewhat fix that with the ability 
to send a lightweight probe and receive lightweight responses over 'UDP' to 
a target host/port. **Supports Windows, macOS and Linux clients.**

## What's Inside The Tin

_The following functions are implemented:_

The core probe function:

- `udp_send_payload`:	Send a UDP payload and return the response (if any)

Helpers for Ubiquity Discover Protocol probes

- `ubnt_discovery_probe`:	Send a Ubiquiti Discovery probe to a host with a discovery query payload and receive a response
udpprobe	Send UDP payloads and gather response
- `parse_ubnt_discovery_response`:	Parser for Ubiquiti Discovery Protocol responses

## Installation

```{r install-ex, eval=FALSE}
devtools::install_git("https://git.sr.ht/~hrbrmstr/udpprobe")
# or
devtools::install_gitlab("hrbrmstr/udpprobe")
# or
devtools::install_github("hrbrmstr/udpprobe")
```

## Usage

```{r lib-ex}
library(udpprobe)

# current version
packageVersion("udpprobe")

```

## Fun with manually crafting DNS records

What's the IP address of `example.com`?

```{r dns}
c(
  0xaa, 0xaa, # id
  0x01, 0x00, # |QR|   Opcode  |AA|TC|RD|RA|   Z    |   RCODE   |
  0x00, 0x01, # QDCOUNT num of questions
  0x00, 0x00, # ANCOUNT 0 since it's only for answers
  0x00, 0x00, # NSCOUNT 0 since it's only for answers
  0x00, 0x00, # ARCOUNT 0 sicne it's only for answers
  0x07,       # start of QNAME beginning with size of "example"
  0x65, 0x78, 0x61, 0x6d, 0x70, 0x6c, 0x65,  # example
  0x03,       # size of "com" (excluding null terminator)
  0x63, 0x6f, 0x6d, 0x00,  # "com" with null terminator
  0x00, 0x01, # QTYPE ("A" == 1)
  0x00, 0x01  # QCLASS (internet == 1)
) -> dns_req

(resp <- udp_send_payload("8.8.8.8", 53, dns_req))

paste0(as.integer(tail(resp, 4)), collapse = ".")

curl::nslookup("example.com")
```

## Live Ubiquiti Test (internet host redacted)

```{r ubnt}
# the following just does:
#   udp_send_payload(Sys.getenv("UBNT_TEST_HOST"), 10001L, c(0x01, 0x00, 0x00, 0x00))
(x <- ubnt_discovery_probe(Sys.getenv("UBNT_TEST_HOST")))

parse_ubnt_discovery_response(x)
```

## udpprobe Metrics

```{r cloc, echo=FALSE}
cloc::cloc_pkg_md()
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.
